---
title: "How to Use the Package"
author: "Hauke Sonnenberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Use the Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7
)
```

## Installation

1. Install the package `remotes` from CRAN
2. Install the package `kwbGompitz` from GitHub

```{r eval = FALSE}
# Install the package "remotes" from CRAN
install.packages("remotes")

# Set the authentification token (ask hauke.sonnenberg@kompetenz-wasser.de)
auth_token <- "1aa89..."

# Install the development version of package "kwbGompitz" from GitHub
devtools::install_github(
  "kwb-r/kwbGompitz@dev", 
  auth_token = auth_token, 
  build_opts = c("--no-resave-data", "--no-manual")
)
```

## GompitZ Model

This package is a wrapper around the model GompitZ by Yves Le Gat. It is 
explained in the Gompitz User's manual. A copy of this manual is contained in
this package. You should be able to open it running the following commands:

```{r eval = FALSE}
path <- "extdata/GompitZ2_08UserGuideOct2011.pdf"

system(paste(options()$pdfviewer, system.file(path, package = "kwbGompitz")))
```

## Usage

### Provide Inspection Data

In the first step you need to provide data on sewer pipe inspections. In the
following I will create a very very small fake dataset from scratch.

The dataset needs to be a data frame with one row per pipe or per inspection. It
is expected to contain the pipe's master data as well as the results of pipe
inspections (if any).

Master data must comprise at least:

- pipe identifier, 
- pipe length, 
- installation year of the pipe and
- name of cohort (= stratum) that the pipe belongs to.

Pipe inspections must be given as pairs of 

- inspection year and
- inspected condition.

If a pipe is not inspected, there is one row for that pipe with the columns
for inspection year and inspected condition left blank (`NA`). 

If a pipe has been inspected, there must be as many rows for that pipe as there
are inspections of that pipe. In the following this is the case for pipe "p1". 
It was inspected twice. 

The pipe's length is expected to be in a column named "LENGTH".

```{r}
inspections <- data.frame(
  cohort = c("main", "main", "main", "main", "main", "main"), 
  pipe_id = c("p1", "p1", "p2", "p3", "p4", "p5"),
  LENGTH = c(50, 50, 60, 55, 58, 53),
  installation_year = as.integer(c(1920, 1920, 1940, 1960, 1930, 1980)),
  inspection_year = as.integer(c(1970, 2010, 2000, 2010, 1995, 2015)),
  condition = c("good", "bad", "good", "good", "bad", "good"),
  stringsAsFactors = FALSE
)

# Just to verify that the fake data look reasonable: calculate the age
inspections$age <- inspections$inspection_year - inspections$installation_year

# Have a look at what we created. Does that look reasonable? Pipes older than 
# 50 years are in bad condition
inspections
```

### View Inspection Data

The package provides a function that aims at getting an overview on the 
inspection data. Try it out:

```{r}
kwbGompitz::plot_pipe_conditions(
  inspections, 
  column_instyear = "installation_year", 
  column_ident = "pipe_id", 
  column_inspyear = "inspection_year", 
  column_condition = "condition", 
  colour_values = c(bad = "red", good = "green")
)

```

### Prepare Model Input Data

```{r}
# Select the columns that are required by the model
masterdata <- kwbGompitz:::composeMasterData(
  stratum = inspections$cohort,
  pipeid = inspections$pipe_id, 
  instyear = inspections$installation_year, 
  inspyear = inspections$inspection_year, 
  condition = inspections$condition
)
```

Create a status matrix, as simple as possible:

1. no covariates
2. length does not influence ageing

See the Gompitz User's Guide as described above.

```{r}
status_matrix <- kwbGompitz::createStatusMatrix(
  strata = unique(inspections$cohort),
  covariates = c(LENGTH = 0)
)

# Compose the input data for the model
input_data <- kwbGompitz::composeGompitzInputData(
  masterdata = masterdata, 
  covariates.status = status_matrix,
  covariates = inspections[, "LENGTH", drop = FALSE],
  weight = 1,
  condition.labels = c("good", "bad")
)
```

### Calibrate the model

```{r}
calibration <- kwbGompitz::runGompitzCalibration(input.data = input_data)

# Plot the survival curves for the only stratum "main"
kwbGompitz::plot_survival_curves(calibration)
```

### Run a prediction

```{r}
prediction <- kwbGompitz::runGompitzPrediction(
  input_data, 
  calibration = calibration, 
  range.years = c(2020, 2070)
)

# Rename column "PredictionYear" to "InspectionYear" 
# (unfortunately the functions to not fit well together...)
prediction_for_plot <- kwb.utils::renameColumns(prediction, list(
  PredictionYear = "InspectionYear"
))

# Plot the prediction
kwbGompitz:::plotPrediction(prediction_for_plot, do_print = FALSE) +
  ggplot2::scale_x_discrete(breaks = seq(2020, 2070, by = 5))
```
